\sysfull

### 2017-08-07 paper/background.tex

\text{Memory Consistency Enforcement Overhead:} Intermittent execution models incur an overhead because they must checkpoint data~\cite{dino,ratchet,quickrecall,mementos}, manage channels~\cite{chain}, or privatize and commit data~\cite{alpaca}. Table~\ref{table:chechpoint_comparison} compares the overheads for several recent intermittent execution models (generalizing discussion from~\cite[Sec. 2.4]{alpaca}.) The mechanism responsible for the overhead of each model varies, but all the bulk of overhead in all models is a manipulating non-volatile memory. Checkpointing moves volatile memory and registers to and from non-volatile memory, channel accesses directly manipulate non-volatile data~\cite{chain}, privatization copies data from and commits data to non-volatile memory~\cite{alpaca}, and idempotence solutions~\cite{ratchet} require all of memory to be non-volatile.

@@

\textbf{Fixed-size Tasks are Inflexible and Inefficient:} A programmer must write each task with statically defined boundaries, which makes statically sized tasks inflexible and non-portable. The programmer must ensure statically that a task's energy requirements match the device's energy storage capacity. If a programmer writes a program that consumes more energy than the device can buffer, the task will never be able to complete using the buffered energy. The consequence of such tasks is a non-terminating program that repeatedly attempts to re-execute the task of excess energy cost. If a programmer writes a program that consumes less energy than the device can buffer, the system may operate inefficiently. When a task and its successor both complete, the two tasks were interleaved by a task boundary that preserves progress and state (i.e., checkpoint or commit).  If the two tasks were not also interleaved by a power failure, then the computation done to preserve state was {\em unnecessary}, only incurring an overhead on the tasks' executions. Avoiding excessively costly, non-terminating tasks and short, high-overhead tasks is a programming challenge, given fixed hardware with a fixed energy buffering capacity that might be \emph{different} for each transiently-powered device (see Table~\ref{table:capacitor}). Such tasks also make it extremely difficult to {\em port} code from  one device to another. A task that is excessively costly on a device with a small energy buffer may become a short, high-overhead task on a device with a larger energy buffer and vice versa. The key problem is that the programmer writes tasks statically for an energy buffer of a fixed size. \sys's {\em task coalescing} mechanism, which we describe in Section~\ref{sec:task_coalescing}, is motivated by the observation that systems with statically defined tasks may be inflexible and inefficient.

>>>>

\noindent \textbf{Memory Consistency Enforcement Overhead:} Intermittent execution models incur an overhead because they must checkpoint data~\cite{dino,ratchet,quickrecall,mementos}, manage channels~\cite{chain}, or privatize and commit data~\cite{alpaca}. Table~\ref{table:chechpoint_comparison} compares the overheads for several recent intermittent execution models (generalizing discussion from~\cite[Sec. 2.4]{alpaca}.) The mechanism responsible for the overhead of each model varies, but all the bulk of overhead in all models is a manipulating non-volatile memory. Checkpointing moves volatile memory and registers to and from non-volatile memory, channel accesses directly manipulate non-volatile data~\cite{chain}, privatization copies data from and commits data to non-volatile memory~\cite{alpaca}, and idempotence solutions~\cite{ratchet} require all of memory to be non-volatile.

@@

\noindent \textbf{Fixed-size Tasks are Inflexible and Inefficient:} A programmer must write each task with statically defined boundaries, which makes statically sized tasks inflexible and non-portable. The programmer must ensure statically that a task's energy requirements match the device's energy storage capacity. If a programmer writes a program that consumes more energy than the device can buffer, the task will never be able to complete using the buffered energy. The consequence of such tasks is a non-terminating program that repeatedly attempts to re-execute the task of excess energy cost. If a programmer writes a program that consumes less energy than the device can buffer, the system may operate inefficiently. When a task and its successor both complete, the two tasks were interleaved by a task boundary that preserves progress and state (i.e., checkpoint or commit).  If the two tasks were not also interleaved by a power failure, then the computation done to preserve state was {\em unnecessary}, only incurring an overhead on the tasks' executions. Avoiding excessively costly, non-terminating tasks and short, high-overhead tasks is a programming challenge, given fixed hardware with a fixed energy buffering capacity that might be \emph{different} for each transiently-powered device (see Table~\ref{table:capacitor}). Such tasks also make it extremely difficult to {\em port} code from  one device to another. A task that is excessively costly on a device with a small energy buffer may become a short, high-overhead task on a device with a larger energy buffer and vice versa. The key problem is that the programmer writes tasks statically for an energy buffer of a fixed size. \sys's {\em task coalescing} mechanism, which we describe in Section~\ref{sec:task_coalescing}, is motivated by the observation that systems with statically defined tasks may be inflexible and inefficient.

### 2017-08-07 paper/compiler.tex
merged heuristic eval section into compiler text

%The compiler consists of two {\em passes} that runs on LLVM Intermediate Representation (IR)\todo{Which version?}{Kiwan}---an intermediate language generated by Clang~\cite{clang_website}:
%
%\begin{enumerate}
%	\item \textbf{Analysis Pass}, which analyzes the code structure and decides where to put task boundaries;
%	\item \textbf{Transform Pass}, which converts the code into task-style program and fills it with necessary \sys keywords so that it can run on \sys.
%\end{enumerate}
%
%We shall now proceed with describing these two parts in detail.

@@

%\sys meets these correctness and efficiency requirements with a four-part
%analysis composed of (i) \emph{graph construction}, (ii) \emph{loop analysis},
%(iii) \emph{task preparation}, and (iv) \emph{single entry assurance}.

@@

The input to the task boundary identification analysis is a standard C program
(as shown in Figure~\ref{fig:comp_1}) and the output of \sys's analysis is the
CFG, with some vertices marked as {\em task boundaries}. 
%
The \sys compiler builds the program's CFG (as shown in
Figure~\ref{fig:comp_2}).  To simplify interprocedural analysis, our prototype
inlines all function calls into the main function in the CFG. 
%
\sys uses a cost-based task boundary placement analysis that
considers per-block energy consumption estimates, inter-task variable liveness,
and the cost of partitioning loops.  We first describe how \sys measures each of these costs 
and then describe \sys's procedure for placing boundaries in the CFG.

\subsubsection{Task Cost Analysis}
\sys's task cost analysis considers estimated task energy cost, variable
liveness, and the impact of splitting loops across tasks.  The result of the
task cost analysis determines where to place boundaries in the CFG.

{\noindent \bf Estimated energy cost.} \sys assigns each basic block in the
CFG a {\em cost}, which is an estimate of the energy consumption of the block.
Precisely modeling the energy cost of arbitrary C code is a difficult problem
and beyond the scope of this work; we used the number of intermediate
representation (IR) instructions in a block as a proxy for energy because
energy cost is likely to closely correspond to block length.  Moreover, we
assume as an input to the compiler the maximum number of operations that the
device can execute without exhausting its buffered energy. This parameter is
highly device-dependent, and we assume that a developer or device manufacturer
determines the number empirically, as we did in this work through direct
measurement.  \sys's analysis tries ensure that the estimated energy cost of different tasks
is roughly the same and does not exceed given energy budget.

@@

After completing its cost analysis of a program, \sys inserts task boundaries
into the CFG in two steps.  The first step is to insert task boundaries on the
back-edges of all loops that have an energy cost greater than the device's
energy buffer capacity, producing an initial boundary placement. The second
step is to insert task boundaries at additional points in the code,
sub-dividing tasks that have an estimated cost that exceeds the energy buffer's
capacity.

@@

The analysis uses a heuristic to compute a score for each candidate and places
a boundary at the candidate with the highest score. The heuristic tries to
avoid placing boundaries on loops, to minimize the number of live variables
that must be protected across the new boundary, and to maximize the number of
operations along the path.  A boundary candidate's score is computed as:
%
\begin{equation}
\text{score} = \frac{w_{0} N_{\text{IR}} - w_{1} N_{\text{PV}}}{N_{\text{loop}}},\nonumber
\end{equation}
%
where $N_{\text{IR}}$ is the length of the path being traversed up to the
candidate, $N_{\text{PV}}$ is the number of variables live variables at the
candidate, and $N_{\text{loop}}$ is the product of the loop bounds of all loops
(i.e., nested loops) containing the boundary candidate. \sys normalizes
$N_{\text{IR}}$ and $N_{\text{PV}}$ to the values among boundary candidates.
The weights, $w_{0}, w_{1}$ are between zero and one, and are
hand-tuned parameters that define the relative importance of $N_{\text{IR}}$,
$N_{\text{PV}}$, and $N_{\text{loop}}$.  We empirically determined these weight
values by measuring the runtime for different weight values across
applications.  We found that varying weights varies runtime by only around 5\%
in most cases and in our experimental evaluation, we used $w_{0} = w_{1} =
0.5$.

After placing task boundaries, the compiler must ensure that all tasks are
single-entry.  To do so, \sys computes the dominator tree of the CFG and checks
that every vertex within a task is dominated by the entry vertex of the task.
If this dominance condition is unmet for a CFG node, the compiler places
an additional task boundary at that node, meeting the condition.
%
Figure~\ref{fig:comp_2} illustrates boundary placement: the first three
vertices make up one task and the remainder become a second task.

@@

After forming tasks, variables that are live beyond the end of the task become
protected, which requires renaming those variables and updating references to
renamed variable names.  
%
For example, Figure~\ref{fig:comp_2} shows a boundary placed within the live
range of {\tt b}; {\tt b} becomes protected and accesses to {\tt b} should be
redirected to {\tt p\_b} (as shown in Figure~\ref{fig:comp_3}).  {\tt a} is
task-local and need not be protected, but only declared in the task. Constants
and read-only variable are never protected.
%
When writes in multiple, newly formed tasks may update the same location (i.e.,
at a PHI node in the CFG), \sys re-writes the accesses to refer to the same global
name.   Subsequent reads simply read the protected variable.

>>>>

@@

%\sys meets these correctness and efficiency requirements with a four-part analysis composed of (i) \emph{graph construction}, (ii) \emph{loop analysis}, (iii) \emph{task preparation}, and (iv) \emph{single entry assurance}.

@@

The input to the task boundary identification analysis is a standard C program (as shown in Figure~\ref{fig:comp_1}) and the output of \sys's analysis is the CFG, with some vertices marked as {\em task boundaries}. The \sys compiler builds the program's CFG (as shown in
Figure~\ref{fig:comp_2}). To simplify inter-procedural analysis, our prototype inlines all function calls into the main function in the CFG. \sys uses a cost-based task boundary placement analysis that considers per-block energy consumption estimates, inter-task variable liveness, and the cost of partitioning loops. We first describe how \sys measures each of these costs and then describe \sys's procedure for placing boundaries in the CFG.

\subsubsection{Task Cost Analysis}

\sys's task cost analysis considers estimated task energy cost, variable liveness, and the impact of splitting loops across tasks. The result of this analysis determines where to place boundaries in the CFG.

{\noindent \bf Estimated energy cost.} \sys assigns each basic block in the CFG a {\em cost}, which is an estimate of the energy consumption of the block. Precisely modelling the energy cost of arbitrary C code is a difficult problem and beyond the scope of this work; we used the number of intermediate representation (IR) instructions in a block as a proxy for energy because energy cost is likely to closely correspond to block length. Moreover, we
assume as an input to the compiler the maximum number of operations that the device can execute without exhausting its buffered energy. This parameter is highly device-dependent, and we assume that a developer or device manufacturer determines the number empirically, as we did in this work through direct measurement. \sys's analysis tries ensure that the estimated energy cost of different tasks is roughly the same and does not exceed given energy budget.

@@

After completing program cost analysis, \sys inserts task boundaries into the CFG in two steps. The first step is to insert task boundaries on the back-edges of all loops that have an energy cost greater than the device's energy buffer capacity, producing an initial boundary placement. The second step is to insert task boundaries at additional points in the code, sub-dividing tasks that have an estimated cost that exceeds the energy buffer's capacity.

@@

The analysis uses a heuristic to compute a score for each candidate and places a boundary at the candidate with the highest score. The heuristic tries to avoid placing boundaries on loops, to minimize the number of live variables that must be protected across the new boundary, and to maximize the number of operations along the path. A boundary candidate's score is computed as
%
\begin{equation}
\text{score} = \frac{w_{0} N_{\text{IR}} - w_{1} N_{\text{PV}}}{N_{\text{loop}}},\nonumber
\end{equation}
%
where $N_{\text{IR}}$ is the length of the path being traversed up to the candidate, $N_{\text{PV}}$ is the number of variables live variables at the candidate, and $N_{\text{loop}}$ is the product of the loop bounds of all loops (i.e., nested loops) containing the boundary candidate. \sys normalizes $N_{\text{IR}}$ and $N_{\text{PV}}$ to the values among boundary candidates. The weights, $w_{0}, w_{1}$ are between zero and one, and are hand-tuned parameters that define the relative importance of $N_{\text{IR}}$,
$N_{\text{PV}}$, and $N_{\text{loop}}$. We empirically determined these weight values by measuring the runtime for different weight values across applications. We found that varying weights varies runtime by only around 5\% in most cases and in our experimental evaluation, we used $w_{0} = w_{1} = 0.5$.

After placing task boundaries, the compiler must ensure that all tasks are single-entry. To do so, \sys computes the dominator tree of the CFG and checks that every vertex within a task is dominated by the entry vertex of the task. If this dominance condition is unmet for a CFG node, the compiler places an additional task boundary at that node, meeting the condition. Figure~\ref{fig:comp_2} illustrates boundary placement: the first three vertices make up one task and the remainder become a second task.

@@

After forming tasks, variables that are live beyond the end of the task become protected, which requires renaming those variables and updating references to renamed variable names. For example, Figure~\ref{fig:comp_2} shows a boundary placed within the live range of {\tt b}; {\tt b} becomes protected and accesses to {\tt b} should be redirected to {\tt p\_b} (as shown in Figure~\ref{fig:comp_3}). {\tt a} is task-local and need not be protected, but only declared in the task. Constants and read-only variable are never protected. When writes in multiple, newly formed tasks may update the same location (i.e., at a PHI node in the CFG), \sys re-writes the accesses to refer to the same global name. Subsequent reads simply read the protected variable.

### 2017-08-07 paper/evaluation.tex

We are now ready to present experimental results of \sys performance. We start with describing the hardware environment (Section~\ref{sec:results_hardware}) followed by the software environment (Section~\ref{sec:results_software}) and we concluded by the evaluation results (Section~\ref{sec:results_evaluation}).

@@

@@

expected results:
\begin{itemize}
	\item performance for each application comparing bars for Chain, no coalescing/hand-tasked, coalescing/hand-tasked, no-coalescing/compiler, coalescing/compiler. Performance depicted, normalized to one of the configurations (probably Chain). This plot will cover most of the eval dimensions that we care about.
	\item Sensitivity plot---show amount of task merged, time to completion, overhead (in terms of task-size/cap-size sweep (remember we can vary either with approximately the same effect)), energy consumed. All experiments for fixed power, at 3 distances with wireless power, compared against Chain for 3 task/cap size configurations
\end{itemize}

>>>>

We present the experimental results of \sys performance. We start with describing the hardware environment (Section~\ref{sec:results_hardware}) followed by the software environment (Section~\ref{sec:results_software}) and we concluded by the evaluation results (Section~\ref{sec:results_evaluation}).

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/virtualTaskSize.eps}
	\caption{Execution time of a dummy application containing twelve empty tasks as a function of coalescing factor. We observe three times improvement from coalescing until four virtual tasks, while we observe a diminishing return from task coalescing beyond four. \todo{Improve the figure: make it narrower, larger font, remove legend, do not capitalize "Task" in X axis, add Y axis label (Execution time (ms))}{Amjad}}
	\label{fig:virtualTaskSize}
\end{figure}

The question remains: does coalescing all program's tasks into a singe virtual task (as the energy becomes fully available, i.e. from energy harvesting to continuous power environment) is the best one?. To answer this question we have implemented a simple program composed of $x$ empty tasks. We execute the same program by increasing coalescing factor by one while the device on which the program was executed was running on a stable power supply (refer to Section~\ref{sec:methodology_evaluation} for details). The result is presented in Figure~\ref{fig:virtualTaskSize}. We observe a significant increase in execution time as task are coalescent from one to four (three times execution time improvement). However, as the coalescing increases beyond four, we observe diminishing returns from coalescing in terms of execution time. \todo{Decide whether to place this figure in Evaluation section or leave it here}{Przemek}

@@

Expected results:
\begin{itemize}
	\item performance for each application comparing bars for Chain, no coalescing/hand-tasked, coalescing/hand-tasked, no-coalescing/compiler, coalescing/compiler. Performance depicted, normalized to one of the configurations (probably Chain). This plot will cover most of the eval dimensions that we care about.
	\item Sensitivity plot---show amount of task merged, time to completion, overhead (in terms of task-size/cap-size sweep (remember we can vary either with approximately the same effect)), energy consumed. All experiments for fixed power, at 3 distances with wireless power, compared against Chain for 3 task/cap size configurations
\end{itemize}
